---
output: github_document
editor_options:
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# staticimports

<!-- badges: start -->
<!-- badges: end -->

**staticimports** provides utility functions that R programmers can use in their packages or other projects. The "static" part means that the functions are statically "imported" -- that is they copied into the project as text, instead of being loaded from a package at run time.

The benefits of doing things this way:

* A package/project that uses staticimports will not take a run-time dependency staticimports.
* Any changes to a function in **staticimports** will not affect a project that uses them, until the author decides to re-import them.
* You can specify the individual functions to import. This is faster and lighter-weight than loading a separate package, especially if the package has much more functionality than you need.

Instead of copying and pasting utility functions from project to project, the utility functions can be centralized in **staticimports**, where they can be vetted and tested.

The functions in **staticimports** are designed to be:

* Fast
* Simple
* Have no external dependencies

If your project imports a function from **staticimports**, and that function changes in a way that has a negative impact on your project, you can simply stop importing it from **staticimports**, and copy the old version to a separate file.


## Installation

You can install the development version of staticimports with:

```{r eval=FALSE}
remotes::install_github("wch/staticimports")
```

## Usage

To use, put a comment block starting with `# @staticimports` in one of your R source files. For example, your `utils.R` may have this at the top:

```
# @staticimports
#  os_name %||%
#  map walk
```

This comment block names the functions to copy from staticimports. To do the import, run:

```{r eval=FALSE}
staticimports::import()
```

By default this will write these funcitons to a file `R/staticimports.R` in your project.

You examine the output by writing to `stdout()` instead of `R/staticimports.R`. Notice how importing `os_name` automatically brings in `is_windows`, `is_mac`, and `is_linux`.

```{r eval=FALSE}
staticimports::import(outfile = stdout())
```

```{r echo=FALSE}
staticimports::import_objs(c("os_name", "%||%", "map", "walk"), stdout())
```

For testing what the output will look like, you can use `import_objs()` to see what it looks like when you import specific objects by name:

```{r}
staticimports::import_objs(c("map", "walk"), stdout())
```


## TODO

* Figure out and explain licensing issues.
* Write license info in generated file.

## Some conventions

Note: This is currently just a dumping ground for ideas.

### Don't use `force()`

Two reasons:

* Error reporting
* Missing args

```{r error=TRUE}
with_option <- function(x, y) {
  force(x)
}

with_option(a = 123)

with_option2 <- function(x, y) {
  x
}

with_option2(a = 123)

err <- function() { stop("asdf") }

f()
```



### Don't use `match.arg()`

* `match.arg()` is slow, at least if values are inferred automatically.
