---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# staticutils

<!-- badges: start -->
<!-- badges: end -->

**staticutils** provides utility functions that R programmers can use in their packages or other projects. The "static" part means that the functions are statically "imported" -- that is they copied into the project as text, instead of being loaded from a package at run time.

The benefits of doing things this way:

* A package/project that uses staticutils will not take a run-time dependency staticutils.
* Any changes to a function in **staticutils** will not affect a project that uses them, until the author decides to re-import them.
* You can specify the individual functions to import. This is faster and lighter-weight than loading a separate package, especially if the package has much more functionality than you need.

Instead of copying and pasting utility functions from project to project, the utility functions can be centralized in **staticutils**, where they can be vetted and tested.

The functions in **staticutils** are designed to be:

* Fast
* Simple
* Have no external dependencies

If your project imports a function from **staticutils**, and that function changes in a way that has a negative impact on your project, you can simply stop importing it from **staticutils**, and copy the old version to a separate file.


## Installation

You can install the development version of staticutils with:

```{r eval=FALSE}
remotes::install_github("wch/staticutils")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r eval=FALSE}
library(staticutils)

staticutils::import()
```

## Some conventions

### Don't use `force()`

Two reasons:

* Error reporting
* Missing args

```{r error=TRUE}
with_option <- function(x, y) {
  force(x)
}

with_option(a = 123)

with_option2 <- function(x, y) {
  x
}

with_option2(a = 123)

err <- function() { stop("asdf") } 

f()
```



### Don't use `match.arg()`

* `match.arg()` is slow, at least if values are inferred automatically.
