---
title: "Speed tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Speed tests}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

format_percent <- function(x) {
  round(x * 100)
}
```

The functions included in staticimports are meant to be fast. For many of the functions, many implementations have been considered. This document contains performance tests of the various possible implementations.

## purrr-like functions

### `walk`

```{r cache=TRUE}
walk_purrr <- purrr::walk

walk_lapply <- function(.x, .f, ...) {
  lapply(.x, .f, ...)
  NULL
}

walk_for <- function(.x, .f, ...) {
  for (i in seq_along(.x)) {
    .f(.x[[i]], ...)
  }
  NULL
}

x <- 1:100
f <- function(a) a
(times <- bench::mark(
  walk_purrr(x, f),
  walk_lapply(x, f),
  walk_for(x, f),
  check = FALSE
))
```


```{r echo=FALSE}
medians <- as.numeric(times$median)
```

Of the three implementations, `walk_for()` is the fastest. It is about `r format_percent(1 - medians[3]/medians[1])`% faster than `walk_purrr()`, and `r format_percent(1 - medians[2]/medians[1])`% faster than `walk_lapply()`.

Note that `walk_purrr()` returns the input `.x` object, whereas the other two implementations return `NULL`.


### `map`

```{r cache=TRUE}
map_purrr <- purrr::map

map_lapply <- function(.x, .f, ...) {
  lapply(.x, .f, ...)
}

map_for <- function(.x, .f, ...) {
  res <- vector("list", length(.x))
  for (i in seq_along(.x)) {
    res[[i]] <- .f(.x[[i]], ...)
  }
  names(res) <- names(.x)
  res
}

x <- 1:100
f <- function(a) a
bench::mark(
  map_purrr(x, f),
  lapply(x, f),     # Bare lapply() for comparison
  map_lapply(x, f),
  map_for(x, f),
)
```

All of these implementations are within a pretty close range. Although `map_for()` is fastest, the margin is small, so we'll just use `map_lapply()` for simplicity.


### `map2`


```{r cache=TRUE}
map2_purrr <- purrr::map2

map2_mapply <- function(.x, .y, .f, ...) {
  mapply(.f, .x, .y, MoreArgs = list(...), SIMPLIFY = FALSE)
}

map2_for <- function(.x, .y, .f, ...) {
  res <- vector("list", length(.x))
  for (i in seq_along(.x)) {
    res[[i]] <- .f(.x[[i]], .y[[i]], ...)
  }
  names(res) <- names(.x)
  res
}

x <- 1:100
y <- x * 1000
f <- function(a, b) a+b
bench::mark(
  map2_purrr(x, y, f),
  map2_mapply(x, y, f),
  map2_for(x, y, f),
)

# With named vector
names(x) <- as.character(x)
(times <- bench::mark(
  map2_purrr(x, y, f),
  map2_mapply(x, y, f),
  map2_for(x, y, f),
))
```

```{r echo=FALSE}
medians <- as.numeric(times$median)
```

`map2_for` is about `r format_percent(1 - medians[3]/medians[2])`% faster than `map2_mapply`, for both named and unnamed inputs, so we'll use `map2_for()`.


### `map2_lgl`, `map2_int`, `map2_dbl`, `map2_chr`

The `map2*` functions return an atomic vector of the specified type.

```{r cache=TRUE}
# Use map2_for implementation from previous section
map2 <- map2_for

map2_dbl_purrr <- purrr::map2_dbl

map2_dbl_mode <- function(.x, .y, .f, ...) {
  res <- map2(.x, .y, .f, ...)
  mode(res) <- "double"
  res
}

map2_dbl_storagemode <- function(.x, .y, .f, ...) {
  res <- map2(.x, .y, .f, ...)
  storage.mode(res) <- "double"
  res
}

# This version is not strictly the same as the others, because it drops names.
map2_dbl_asnumeric <- function(.x, .y, .f, ...) {
  res <- as.numeric(map2(.x, .y, .f, ...))
  names(res) <- names(.x)
  res
}

map2_dbl_for <- function(.x, .y, .f, ...) {
  res <- vector("double", length(.x))
  for (i in seq_along(.x)) {
    res[[i]] <- .f(.x[[i]], .y[[i]], ...)
  }
  names(res) <- names(.x)
  res
}

x <- 1:1000
y <- x * 10000
x <- as.list(x)
y <- as.list(y)
f <- function(a, b) a+b

# This is what the output should look like
map2_dbl_purrr(1:3, 101:103, f)

bench::mark(
  map2_dbl_purrr(x, y, f),
  map2_dbl_mode(x, y, f),
  map2_dbl_storagemode(x, y, f),
  map2_dbl_asnumeric(x, y, f),
  map2_dbl_for(x, y, f),
)

# Same test, with names
names(x) <- as.character(x)
(times <- bench::mark(
  map2_dbl_purrr(x, y, f),
  map2_dbl_mode(x, y, f),
  map2_dbl_storagemode(x, y, f),
  map2_dbl_asnumeric(x, y, f),
  map2_dbl_for(x, y, f),
))
```

```{r echo=FALSE}
medians <- as.numeric(times$median)
```

`map2_dbl_for()` is the fastest by a bit. However, one drawback is that if `.f()` returns a value of the incorrect type, it simply promotes the result vector to that type, and emits no warnings; the returned vector is not guaranteed to be of the specified type. This is not acceptable behavior.

```{r}
x <- c(1, 2)
map2_dbl_for(x, x, function(a, b) "test")
```

The ideal behavior in this situation is for the function to throw an error when an incorrect type is returned. This is what `map2_dbl_purrr()` does.

```{r error=TRUE}
map2_dbl_purrr(x, x, function(a, b) "test")
```

The other three versions, `map2_dbl_mode()`, `map2_dbl_storagemode()`, and `map2_dbl_asnumeric()` emit warnings, which isn't ideal but it is acceptable.

```{r error=TRUE}
map2_dbl_mode(x, x, function(a, b) "test")
map2_dbl_storagemode(x, x, function(a, b) "test")
map2_dbl_asnumeric(x, x, function(a, b) "test")
```

`map2_dbl_asnumeric()` has the best balance of speed, understandability, and warning/error behavior when the function returns the wrong type.
